Endogenous Politics and the Design of Trade Institutions\thanks{I thank Kyle Bagwell, Mostafa Beshkar, Rick Bond, Chad Bown, Nuno Limao, Giovanni Maggi, Anson Soderbery, Bob Staiger, Maurizio Zanardi, Ben Zissimos and participants in seminars at SITE for very helpful comments and discussions.}

### 2017-01-16 draft.tex
cannibalize incentive compatibility subsection under model

In the first stage, the governments choose the trade agreement tariffs via a negotiating process that I assume to be efficient. This process therefore maximizes the joint payoffs produced by the trade agreement:
\begin{equation}
  W(\ga(s,e),\ga^*(s^*,e^*),\tau,\tau^*) = W(\ga(s,e),\tau,\tau^*) + W^*(\ga^*(s^*,e^*),\tau,\tau^*)
  \label{eq:jv}
\end{equation}

Government welfare in both countries is additively separable in $\tau$ and we can write the welfare of the foreign government as $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$. Therefore maximizing Expression~\ref{eq:jv} in $\tau$ is equivalent to maximizing $W_X(\ga(s,e),\tau) + W_X^*(\tau)$. The trade agreement tariff is then given by
\begin{equation}
  \tau^a(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau) + W_X^*(\tau)
  \label{eq:eff}
\end{equation}
As in the applied tariff problem, I assume the realization of $\ga$ is below the threshold need to ensure the second order condition is satisfied.

@@

\subsection{Exogenous Political Pressure and Incentive Compatibility}
In order to compare the results for the case of endogenous political pressure to \Textcite{bs2005} results for exogenous political shocks, I provide the essentials of their setup.

They assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$. The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive. They also assume $\ga$ and $\ga^*$ are private information.

This leads to a concern about incentive compatibility. It must be in a government's best interest to truthfully reveal its $\ga$ or else it will misrepresent its private information about $\ga$ in order to raise tariffs and create a terms-of-trade gain, thereby improving its unilateral welfare. They provide conditions that guarantee incentive compatibility on page 481.

On a separate note, it is important to make clear that in this strand of the literature, it is assumed that trade agreements are negotiated to maximize the expected welfare of politically motivated governments given the exogenously-determined political pressure they expect to face at the time of the agreement's implementation. I will follow this convention throughout.

@@

This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.\footnote{Note that in their treatment, Bagwell and Staiger (2005) assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.

@@

However, because the tariff allowed under the trade agreement with an escape clause varies with the announced level of political pressure, the incentive compatibility constraint now comes into play: a government may gain an advantage by misreporting the level of $\ga$ it experiences. In their Proposition 4, Bagwell and Staiger show that a trade agreement with a costless escape clause cannot improve welfare because it cannot be made incentive compatible. No matter the realization of the stochastic $\ga$, it is always in the government's interest to announce that $\ga$ is high, which allows it to apply any tariff up to the higher weak binding. This deviation improves unilateral welfare while imposing a terms-of-trade externality on the trading partner.

>>>>

In the first stage, the governments choose the trade agreement tariffs via a negotiating process that I assume to be efficient. In this strand of the literature, it is typically assumed that trade agreements are negotiated to maximize the expected welfare of politically motivated governments given the exogenously-determined political pressure they expect to face at the time of the agreement's implementation. I will follow this convention throughout, making necessary adjustments for endogenous politics. The trade negotiation therefore is assumed to maximize the joint payoffs produced by the trade agreement:
\begin{equation}
  W(\ga(s,e),\ga^*(s^*,e^*),\tau,\tau^*) = W(\ga(s,e),\tau,\tau^*) + W^*(\ga^*(s^*,e^*),\tau,\tau^*)
  \label{eq:jv}
\end{equation}

Government welfare in both countries is additively separable in $\tau$ and we can write the welfare of the foreign government as $W^*(\ga^*(s^*,e^*),\tau,\tau^*)$ = $W_Y^*(\ga^*(s^*,e^*),\tau^*) + W_X^*(\tau)$. Therefore maximizing Expression~\ref{eq:jv} in $\tau$ is equivalent to maximizing $W_X(\ga(s,e),\tau) + W_X^*(\tau)$. The trade agreement tariff is then given by
\begin{equation}
  \tau^a(\ga(s,e)) = \max_\tau W_X(\ga(s,e),\tau) + W_X^*(\tau)
  \label{eq:eff}
\end{equation}
As in the applied tariff problem, I assume the realization of $\ga$ is below the threshold needed to ensure the second order condition is satisfied.

@@

@@

This is in contrast to the case of exogenous political pressure, for which \Textcite{bs2005} show several interesting results.\footnote{\Textcite{bs2005} assume that $\ga$ and $\ga^*$ are each drawn independently from probability distributions with cumulative distribution functions $H(\ga)$ with $h(\ga)=H'(\ga)$. The support for this probability distribution is $\left[\un{\ga}, \ov{\ga} \right]$ where $\ov{\ga}$ is small enough that the government's optimal tariff will be non-prohibitive. They also assume $\ga$ and $\ga^*$ are private information. They also assume linear supply and demand.} In particular, they find that when governments negotiate commitments that take the form of weak bindings, the tariff caps they choose imply that applied tariffs will be set strictly below the bound level when governments experience low realizations of political pressure.

@@

However, because the tariff allowed under the trade agreement with an escape clause varies with the announced level of political pressure, introducing an escape clause leads to a concern about incentive compatibility. It must be in a government's best interest to truthfully reveal its $\ga$ or else it will misrepresent its private information about $\ga$ in order to raise tariffs and create a terms-of-trade gain, thereby improving its unilateral welfare. \Textcite{bs2005} provide conditions that guarantee incentive compatibility, but they also show that a trade agreement with a costless escape clause cannot improve welfare because it cannot be made incentive compatible. No matter the realization of the stochastic $\ga$, it is always in the government's interest to announce that $\ga$ is high, which allows it to apply any tariff up to the higher weak binding. This deviation improves unilateral welfare while imposing a terms-of-trade externality on the trading partner.
